A Constructive Ordered Predicate Logic for Category Theory

### 2022-10-07 paper/paper.tex
cleanup

Now that we have a category-theoretic notion of model, we give some
model construction theorems that can be used to justify our intuitive
notion of semantics in (enriched, internal, indexed) category theory.

First, we can extend any virtual equipment to a hyperdoctrine of
virtual equipments

First, we can extend any model of \ohol{} to a model of \ohol{}
parameterized by dependent type theory as follows:
\begin{construction}
  Given a virtual equipment $\mathcal V$, we
  can construct a functor $\mathcal V^{-}: \Set \to \text{vDbl}_r$ by
  defining the objects of $\mathcal V^\Gamma$ to be functions $\Gamma
  \to \mathcal V_o$, and similarly for morphisms and 2-cells with all
  operations given pointwise. 
\end{construction}
% NOTE: this should be some kind of general theorem right?
% Sketch:
% 1. VDCrs are a finite limit theory
% 2. A hyperdoctrine L : C -> VDCr is the same as a VDCr in Pr(C)
% 3. The Yoneda embedding preserves finite limits

@@

More generally, categories internal to, enriched in and/or indexed by
sufficiently nice categories define a virtual equipment that model
most of the connectives of our type theory. We highlight one example
from the literature that is highly general: Shulman's enriched indexed
categories. Shulman's construction defines a virtual double categories
of large and small $\mathcal V$-categories for any pseudofunctor
$\mathcal V : S^o \to \MonCat$ where $S$ is a category with finite
products. He gives examples that show that this internal, enriched and
indexed categories for suitable choices of $\mathcal V$.
%
This is slightly weaker then what we require: to have
\emph{split} restrictions, we need that $\mathcal V$ be a
\emph{strict} functor, not merely a pseudo-functor.
%
This is analogous to the situation for dependent type theory where
substitution is strictly associative but in models is typically
modeled by pullback which is only associative up to unique
isomorphism. In fact, one of the main examples is only a
pseudo-functor for precisely this reason that the action on morphisms
is given by pullback.
%
We expect this difference can be handled in the same way: by a general
strictification theorem or by moving to a univalent setting.

>>>>

Now that we have a category-theoretic notion of model, we give some
model construction theorems that can be used to justify our intuitive
notion of semantics in (enriched, internal, indexed) category theory.
%
First, we can extend any virtual equipment to an HVE as follows:
\begin{construction}
  Given a virtual equipment $\mathcal V$, we
  can construct a functor $\mathcal V^{-}: \Set \to \text{vDbl}_r$ by
  defining the objects of $\mathcal V^\Gamma$ to be functions $\Gamma
  \to \mathcal V_o$, and similarly for morphisms and 2-cells with all
  operations given pointwise. 
\end{construction}
% NOTE: this should be some kind of general theorem right?
% Sketch:
% 1. VDCrs are a finite limit theory
% 2. A hyperdoctrine L : C -> VDCr is the same as a VDCr in Pr(C)
% 3. The Yoneda embedding preserves finite limits

@@

More generally, categories internal to, enriched in and/or indexed by
sufficiently nice categories define a virtual equipment that model the
connectives of VETT. We highlight one example from the literature that
is highly general: Shulman's enriched indexed categories. Shulman's
construction defines a virtual double categories of large and small
$\mathcal V$-categories for any pseudofunctor $\mathcal V : S^o \to
\MonCat$ where $S$ is a category with finite products. He gives
examples that show that this subsumes ordinary internal, enriched and
indexed categories for suitable choices of $\mathcal V$, as well as
more general categories that can be thought of as both indexed and
enriched.
%
This is slightly weaker then what we require: to have
\emph{split} restrictions, we need that $\mathcal V$ be a
\emph{strict} functor, not merely a pseudo-functor.
%
This is analogous to the situation for dependent type theory where
substitution is strictly associative but in models is typically
modeled by pullback which is only associative up to unique
isomorphism. In fact, one of the main examples is only a
pseudo-functor for precisely this reason that the action on morphisms
is given by pullback.
%
We expect this difference can be handled in the same way: by a general
strictification theorem or by moving to a univalent setting.
